---
title: "Macropodoid Modelling"
author: "Ian G. Brennan"
date: "29/07/2019"
output:
  #html_document:
  #  toc: yes
  pdf_document:
    toc: yes
---

# Getting the Data Together

```{r echo=FALSE, message=FALSE}
library(RPANDA); library(randomcoloR)
library(dplyr); library(gridExtra)
library(geiger); library(phytools)
library(parallel); library(pspline)
library(mvMORPH); library(cladeMode)
library(awtools); library(ggplot2); 
library(ggridges); library(gridExtra)
library(RColorBrewer); library(deeptime);
library(Rmisc); library(scales)


# show_col(mpalette); or pie(rep(1, length(mpalette)), col = mpalette, labels = mpalette)
```

Now read in the tree files we'll be working with 
```{r}
#min.tree <- read.nexus("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/FossilUncertainty/Macro_MinAges_CON.tre")
#mean.tree <- read.nexus("/PATH/Macro_MeanAges_CON.tre")
#max.tree <- read.nexus("/PATH/Macro_MaxAges_CON.tre")
#cp.min <- read.nexus("/PATH/Macro_CP_MinAges_CON.tre")
#cp.mean <- read.nexus("/PATH/Macro_CP_EstAges_CON.tre")
#cp.max <- read.nexus("/PATH/Macro_CP_MaxAges_CON.tre")
sampled.trees   <- read.tree("/Users/Ian/Google.Drive/ANU/Macropod_Dating/REAL_Run5_AllSchemes_508trees_Macropodinae.trees")
empirical.trees <- read.tree("/Users/Ian/Google.Drive/ANU/Macropod_Dating/MODEL110_Sampled_Run2_Macropodinae.trees")
fossil.trees    <- read.tree("/Users/Ian/Google.Drive/ANU/Macropod_Dating/REAL_Run4_Fossil_519trees_Macropodinae.trees")
consensus.tree  <- read.nexus("/Users/Ian/Google.Drive/ANU/Macropod_Dating/Operators/REAL_Macropodinae_ALL/Run5_Combined123_CON.tre")
```

Choose the current tree we want to work with
```{r}
tree <- consensus.tree
```

And the hypsodonty data
```{r}
# raw data for all the macropodoid taxa
all.HI           <- read.csv("/Users/Ian/Google.Drive/ANU/Macropod_Dating/FossilUncertainty/Data/CrownHeight_Data.csv", header=T)
# or just species means for the Macropodinae
hypsodonty.index <- read.csv("/Users/Ian/Google.Drive/ANU/Macropod_Dating/FossilUncertainty/Data/Macropodinae_Hypsodonty_Data.csv", header=T)
head(hypsodonty.index)
```

The C4 plant reconstruction data from Andrae
```{r}
enviro.data <- read.csv("/Users/Ian/Google.Drive/ANU/Macropod_Dating/Andrae_S1.csv", header=T)
  grass.data <- enviro.data[1:25, c(1,3)] 
  # this will also remove the last sample (from 9.5 mya) which is spurious, and keep just the mean estimate
head(enviro.data); head(grass.data)
```
And the dust flux data from Andrae
```{r}
flux.data <- read.csv("/Users/Ian/Google.Drive/ANU/Macropod_Dating/Aeolian_Flux.csv", header=T)
head(flux.data)
```
As well as the paleotemperature data
```{r}
data(InfTemp)
head(InfTemp)
```

Trim tree and data down to overlapping taxa
```{r}
# extract the taxa that are in both the tree and 
overlaps <- intersect(tree$tip.label, unique(hypsodonty.index$Taxon))
macro.tree <- drop.tip(tree, setdiff(tree$tip.label, overlaps))
#macro.tree <- lapply(tree, drop.tip, tip=tip.drops) # if you're using a set of trees (fossil, sampled)
trim.data <- dplyr::filter(hypsodonty.index, Taxon %in% overlaps)
macro.HI <- trim.data[,2]; names(macro.HI) <- trim.data[,1]; geiger::name.check(macro.tree, macro.HI)
macro.HI
```

If you're working with the raw data, trim tree and data down to just Macropodinae
```{r, eval=FALSE}
macros <- dplyr::filter(all.HI, Higher_tax == "Macropodinae")
overlaps <- intersect(tree$tip.label, unique(all.HI$Taxon))
trim.tree <- drop.tip(tree, setdiff(tree$tip.label, overlaps))
trim.raw <- filter(macros, Taxon %in% overlaps)
```

create a tibble to get the species means (if you haven't done this already)
```{r, eval=FALSE}
  library(dplyr)
  sp.means <- trim.raw %>%
    group_by(Taxon) %>%
    summarise_at(vars(H_HYPCD/PW), mean)
  #write.csv(sp.means, row.names=FALSE, file="/PATH/CrownHeight_Macropodinae_spMEANS.csv") # uncomment to write the file
```


# Visualizing Our Data

Let's quickly visualize the data in a few different ways to get an idea of what's going on
```{r}
# Barplot of trait value
plotTree.barplot(macro.tree, macro.HI, args.barplot=list(beside=TRUE, border=F))
```
```{r}
# Continuous trait map
obj1 <- contMap(macro.tree, macro.HI, plot=FALSE, outline=F); 
    n<-length(obj1$cols); 
    obj1$cols[1:n] <- rev(colorRampPalette(brewer.pal(9, "YlGnBu"))(n));
plot(obj1,legend=0.7*max(nodeHeights(obj1$tree)),
     fsize=c(0.7,0.9), lwd=5, border=F); axisPhylo(1, backward=T)
```
```{r}
# Hypsodonty distributed across feeding guilds
hidata <- hypsodonty.index[complete.cases(hypsodonty.index[,c(1,2,4)]),]
    hidata <- hidata[order(hidata$HI),]
        hidata$Taxon <- factor(hidata$Taxon, levels = hidata$Taxon)
ggplot(hidata, aes(x=Taxon, y=HI, fill=Diet_Guild), colour=brewer.pal(3,"Paired")) + 
  geom_col() + theme_classic() + coord_flip() +
  scale_x_discrete(limits = rev(hidata$Taxon)) +
  scale_fill_brewer(palette="Paired", "Diet_Guild")
```

Now we can look at the time-sampled variables
```{r}
# make a plot of the grass data
pp <- ggplot(enviro.data[1:25,], aes(Age)) +
  geom_ribbon(aes(ymin = C4_recon_lower, ymax = C4_recon_upper), fill = "light green") +
  geom_line(aes(y = C4_recon_mean), color="DarkGreen") + scale_x_reverse() + theme_classic() +
  coord_cartesian(xlim = c(0, 20), ylim = c(0,80), expand = FALSE) 
#qq <- gggeo_scale(pp, dat="epochs") # if you want to plot a geological timescale

# make a plot of the flux data
rr <- ggplot(flux.data, aes(Age)) +
  geom_ribbon(aes(ymin = A_Flux-35, ymax = A_Flux+35), fill = "light blue") +
  geom_line(aes(y = A_Flux), color="DarkBlue") + scale_x_reverse() + theme_classic() +
  coord_cartesian(xlim = c(0, 20), ylim = c(0,150), expand = FALSE) 
#ss <- gggeo_scale(rr, dat="epochs") # if you want to plot a geological timescale

grid.arrange(pp, rr, nrow=1)
```


# Fitting Models of Trait Evolution

Correlative models like the environmental models in RPANDA will be sensitive to the amount of smoothing to the trend line of the input data (see Clavel & Morlon, PNAS). To address this, we'll create a function that searches for the optimum smoothness of the trend by fitting a set of values. 
```{r}
best.smoothing <- function (phy, trait.data, time.data=InfTemp, degrees=c(0,10,20,30,40,50), model="EnvExp", cores=6) {
  res.list <- mclapply(1:length(degrees), function(x) {
    fit_t_env(phy, trait.data, env_data=time.data, df=degrees[x], scale=F, plot=T, model=model)}, mc.cores = cores)
  for(i in 1:length(res.list)){res.list[[i]]$df <- degrees[i]}
  res.values <- unlist(lapply(res.list, function(x) x$aicc)) # make a vector of the values, so we can get the index number of the best
  best.res <- res.list[[which.min(res.values)]]
  
  plot(best.res, main=paste(model, "; AICc = ", round(best.res$aicc,2)), sub=paste("sigma = ",round(best.res$param[1],2), " beta = ",round(best.res$param[2],2), " df = ", best.res$df), col="red")
  
  return(list(all.results=res.list, best.result=best.res, best.df=best.res$df))
}
```

We can have a look at what this smoothing actually does to our data. We can come back to look at these once we get the optimum fits for our models.
```{r}
grass.spline0  <- sm.spline(x=grass.data$Age, y=grass.data$C4_recon_mean, df=0)
grass.spline10 <- sm.spline(x=grass.data$Age, y=grass.data$C4_recon_mean, df=10)
grass.spline20 <- sm.spline(x=grass.data$Age, y=grass.data$C4_recon_mean, df=20)
grass.spline30 <- sm.spline(x=grass.data$Age, y=grass.data$C4_recon_mean, df=30)
grass.spline40 <- sm.spline(x=grass.data$Age, y=grass.data$C4_recon_mean, df=40)
grass.spline50 <- sm.spline(x=grass.data$Age, y=grass.data$C4_recon_mean, df=50)

plot(grass.data, main="C4 Grass Reconstruction Through Time", xlab="Millions of Years Ago", ylab="Percent C4")
lines(grass.spline0, col="red", lwd=4)
lines(grass.spline10, col="green", lwd=4)
lines(grass.spline20, col="blue", lwd=4)
lines(grass.spline30, col="yellow", lwd=4)
lines(grass.spline40, col="violet", lwd=4)
lines(grass.spline50, col="gold", lwd=4)

flux.spline0  <- sm.spline(x=flux.data$Age, y=flux.data$A_Flux, df=0)
flux.spline10 <- sm.spline(x=flux.data$Age, y=flux.data$A_Flux, df=10)
flux.spline20 <- sm.spline(x=flux.data$Age, y=flux.data$A_Flux, df=20)
flux.spline30 <- sm.spline(x=flux.data$Age, y=flux.data$A_Flux, df=30)
flux.spline40 <- sm.spline(x=flux.data$Age, y=flux.data$A_Flux, df=40)
flux.spline50 <- sm.spline(x=flux.data$Age, y=flux.data$A_Flux, df=50)

plot(flux.data, main="Aeolian Dust Flux Through Time", xlab="Millions of Years Ago", ylab="Aeolian Dust Flux")
lines(flux.spline0, col="red", lwd=4)
lines(flux.spline10, col="green", lwd=4)
lines(flux.spline20, col="blue", lwd=4)
lines(flux.spline30, col="yellow", lwd=4)
lines(flux.spline40, col="magenta", lwd=4)
lines(flux.spline50, col="cyan", lwd=4)

data(InfTemp)
env.spline0 <- sm.spline(x=InfTemp$Age, y=InfTemp$Temperature, df=0)
env.spline10 <- sm.spline(x=InfTemp$Age, y=InfTemp$Temperature, df=10)
env.spline20 <- sm.spline(x=InfTemp$Age, y=InfTemp$Temperature, df=20)
env.spline30 <- sm.spline(x=InfTemp$Age, y=InfTemp$Temperature, df=30)
env.spline40 <- sm.spline(x=InfTemp$Age, y=InfTemp$Temperature, df=40)
env.spline50 <- sm.spline(x=InfTemp$Age, y=InfTemp$Temperature, df=50)

plot(InfTemp, main="Paleotemperature Through Time")
lines(env.spline0, col="red")
lines(env.spline10, col="green")
lines(env.spline20, col="blue")
lines(env.spline30, col="yellow")
lines(env.spline40, col="magenta")
lines(env.spline50, col="cyan")
```

### Next we'll fit a number of models to our tree and data
Start with the environmental model of paleotemperature. You can designate the number of cores and the amount of smoothing
```{r}
# linear model first
ENVexp <- best.smoothing(macro.tree, macro.HI, time.data=InfTemp, 
                         degrees=c(10,20,30,40,50), model="EnvExp", cores=5)
# exponential model next
ENVlin <- best.smoothing(macro.tree, macro.HI, time.data=InfTemp, 
                         degrees=c(10,20,30,40,50), model="EnvLin", cores=5)
```

Next up the grass model, using C4 reconstructions.
```{r}
GRASSexp <- best.smoothing(macro.tree, macro.HI, time.data=grass.data, 
                           degrees=c(10,20,30,40,50), model="EnvExp", cores=5)
GRASSlin <- best.smoothing(macro.tree, macro.HI, time.data=grass.data, 
                           degrees=c(10,20,30,40,50), model="EnvLin", cores=5)
```

And finally the flux models, using aeolian dust measurements.
```{r}
FLUXexp <- best.smoothing(macro.tree, macro.HI, time.data=flux.data, 
                          degrees=c(10,20,30,40,50), model="EnvExp", cores=5)
FLUXlin <- best.smoothing(macro.tree, macro.HI, time.data=flux.data, 
                          degrees=c(10,20,30,40,50), model="EnvLin", cores=5)
```

Lastly, for comparison, run a few standard models. These are Brownian Motion, Brownian Motion with a Trend, and Early Burst.
```{r}
BM_res <-    fitContinuous(macro.tree, macro.HI, model="BM")
trend_res <- fitContinuous(macro.tree, macro.HI, model="trend")
EB_res <-    fitContinuous(macro.tree, macro.HI, model="EB")
```

Compare the models with AICc, and check differences across the trees
```{r}
model_FIT <- c(ENVexp$best.result$aicc, ENVlin$best.result$aicc, 
               GRASSlin$best.result$aicc, GRASSexp$best.result$aicc, 
               FLUXexp$best.result$aicc, FLUXlin$best.result$aicc, 
               BM_res$opt$aicc, trend_res$opt$aicc, EB_res$opt$aicc); 
names(model_FIT) <- c("ENVexp", "ENVlin", "GRASSlin", "GRASSexp", "FLUXexp", "FLUXlin", "BM", "Trend", "EB")
aic.w(model_FIT)
fit.aic <- as.data.frame(as.vector(aic.w(model_FIT))); fit.aic$model <- names(model_FIT); colnames(fit.aic) <- c("aiccw", "model"); fit.aic$age <- "Consensus Tree"
```
Quickly collapse models from the same data 
```{r}
fit.aic$model.type <- c("ENV", "ENV", "GRASS", "GRASS", "FLUX", "FLUX", "BM", "Trend", "EB")
fit.aic$model.type <- factor(fit.aic$model.type, levels=c("ENV", "FLUX", "GRASS", "BM", "EB", "Trend"))
```
Then plot the model fits as AICc weights
```{r}
library(wesanderson)
(ggplot(fit.aic)
  + geom_bar(aes(y=aiccw, x=age, fill=model.type), stat="identity")
  #+ theme(axis.text.x=element_text(angle=25, hjust=1), panel.background=element_blank(), legend.position="bottom")
  + theme_classic()
  + scale_fill_manual( values=wes_palette("Zissou1", 6, "continuous")))
```


# Fitting Models to Our Data as a Function of Time

Ok, now that we've fit the models to a given tree, we want to fit the models to lots of trees of different ages, shapes, etc. Now we'll run a loop across all of these trees, to fit the models to each one. It may take a little while.
```{r, eval=FALSE}
# MAKE THIS EVAL=TRUE IF YOU WANT TO DO THIS BIT FOR REAL
tree.span <- fossil.trees
mean.data <- fossil.HI
# Fit all the models to a series of trees!
all.aics <- NULL; all.results <- NULL; timer <- progress_estimated(length(tree.span))
for (k in 1:length(tree.span)){
  int.results <- NULL
  
  # Fit a number of models to the data (ENV, GRASS, BM, EB, Trend, Drift)
  ENVexp <- best.smoothing(tree.span[[k]], mean.data, time.data=InfTemp, 
                           degrees=c(10,20,30,40,50), model="EnvExp", cores=5); 
                           int.results[["ENVexp"]] <- ENVexp$best.result; 
  ENVlin <- best.smoothing(tree.span[[k]], mean.data, time.data=InfTemp, 
                           degrees=c(10,20,30,40,50), model="EnvLin", cores=5); 
                           int.results[["ENVlin"]] <- ENVlin$best.result; 
  
  GRASSexp <- best.smoothing(tree.span[[k]], mean.data, time.data=grass.data, 
                             degrees=c(30,40,50), model="EnvExp", cores=3); 
                             int.results[["GRASSexp"]] <- GRASSexp$best.result; 
  GRASSlin <- best.smoothing(tree.span[[k]], mean.data, time.data=grass.data, 
                             degrees=c(30,40,50), model="EnvLin", cores=3); 
                             int.results[["GRASSlin"]] <- GRASSlin$best.result;
  
  FLUXexp <- best.smoothing(tree.span[[k]], mean.data, time.data=flux.data, 
                            degrees=c(10,20,30,40,50), model="EnvExp", cores=5); 
                            int.results[["FLUXexp"]] <- FLUXexp$best.result; 
  FLUXlin <- best.smoothing(tree.span[[k]], mean.data, time.data=flux.data, 
                            degrees=c(10,20,30,40,50), model="EnvLin", cores=5); 
                            int.results[["FLUXlin"]] <- FLUXlin$best.result; 
  
  BM_res    <- fitContinuous(tree.span[[k]], mean.data, model="BM"); 
               int.results[["BM"]] <- BM_res
  trend_res <- fitContinuous(tree.span[[k]], mean.data, model="trend"); 
               int.results[["Trend"]] <- trend_res
  EB_res    <- fitContinuous(tree.span[[k]], mean.data, model="EB"); 
               int.results[["EB"]] <- EB_res
  
  curr_tree_FIT <- c(ENVexp$best.result$aicc, ENVlin$best.result$aicc, 
                    GRASSexp$best.result$aicc, GRASSlin$best.result$aicc, 
                    FLUXexp$best.result$aicc, FLUXlin$best.result$aicc,
                    BM_res$opt$aicc, trend_res$opt$aicc, EB_res$opt$aicc); 
  names(curr_tree_FIT) <- c("ENVexp", "ENVlin", "GRASSexp", "GRASSlin", 
                            "FLUXexp", "FLUXlin", "BM", "Trend", "EB")
  
  curr_tree_SIG <- c(ENVexp$best.result$param[[1]], ENVlin$best.result$param[[1]], 
                     GRASSexp$best.result$param[[1]], GRASSlin$best.result$param[[1]], 
                     FLUXexp$best.result$param[[1]], FLUXlin$best.result$param[[1]],
                     BM_res$opt$sigsq, trend_res$opt$sigsq, EB_res$opt$sigsq); 
  names(curr_tree_SIG) <- c("ENVexp", "ENVlin", "GRASSexp", "GRASSlin", 
                            "FLUXexp", "FLUXlin", "BM", "Trend", "EB")
  
  curr_tree_PAR <- c(ENVexp$best.result$param[[2]], ENVlin$best.result$param[[2]], 
                     GRASSexp$best.result$param[[2]], GRASSlin$best.result$param[[2]], 
                     FLUXexp$best.result$param[[2]], FLUXlin$best.result$param[[2]],
                     NA, trend_res$opt$slope, EB_res$opt$a); 
  names(curr_tree_PAR) <- c("ENVexp", "ENVlin", "GRASSexp", "GRASSlin", 
                            "FLUXexp", "FLUXlin", "BM", "Trend", "EB")
  
  curr.aic <- as.data.frame(as.vector(aic.w(curr_tree_FIT))); 
        curr.aic$model <- names(curr_tree_FIT); colnames(curr.aic) <- c("aiccw", "model"); 
            curr.aic$age <- round(max(nodeHeights(tree.span[[k]])), 4)
                curr.aic$tree <- k
                    curr.aic$sigsq <- curr_tree_SIG
                        curr.aic$par <- curr_tree_PAR
  all.results[[k]] <- int.results
  
  curr.aic$model.type <- c("ENV", "ENV", "GRASS", "GRASS", 
                           "FLUX", "FLUX", "BM", "Trend", "EB")
  
  all.aics <- rbind.data.frame(all.aics, curr.aic); 

  print(timer$tick())
  
}
```

Save the file externally:
```{r, eval=FALSE}
saveRDS(all.aics,    file="/PATH/Fossil_Trees_Model_Fitting_AICCs.RDS")
saveRDS(all.results, file="/PATH/Fossil_Trees_Model_Fitting_Results.RDS")
#saveRDS(all.aics, file="/PATH/Model_Fitting_AICCs.RDS")
```
Or skip the work and read in the file instead:
```{r}
all.aics <- readRDS("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/FossilUncertainty/Data/FINAL_SAMPLED_Model_Fitting_AICCs.RDS")
all.aics$model.type <- factor(all.aics$model.type, levels=c("ENV", "FLUX", "GRASS", "BM", "EB", "Trend"))
all.aics$age <- as.factor(all.aics$age)
```

```{r}
sampled.res <- (ggplot(all.aics)
  + geom_bar(aes(y=aiccw, x=age, fill=model.type), stat="identity")
  + theme(axis.text.x=element_text(angle=90, hjust=1), panel.background=element_blank(), legend.position="none")
  + scale_fill_manual( values=wes_palette("Zissou1", 6, "continuous")))
sampled.res
```

And then summarize the model support (average AICcWt) across all trees
```{r}
# Summarize the model support (average AICcWt)
aic.sum <- summarySE(all.aics, measurevar="aiccw", groupvars="model")
aic.sum$model <- factor(aic.sum$model, levels=c("ENVlin", "ENVexp", "FLUXlin", "FLUXexp", "GRASSexp", "GRASSlin", "BM", "EB", "Trend"))
aic.sum$model.type <- c("BM", "EB", "ENV", "ENV", "FLUX", "FLUX", "GRASS", "GRASS", "Trend")
aic.sum$model.type <- factor(aic.sum$model.type, levels=c("ENV", "FLUX", "GRASS", "BM", "EB", "Trend"))

fossil.bar <- (ggplot(aic.sum, aes(x=model.type, y=aiccw, fill=model.type))
  + geom_bar(stat="identity")
  #+ geom_errorbar(aes(ymin=aiccw-se, ymax=aiccw+se), size=0.3, width=0.2)
  + scale_fill_manual( values=wes_palette("Zissou1", 6, "continuous"))
  + theme_classic()
  + theme(legend.position="none")
  #+ facet_wrap(~group, nrow=3, ncol=2)
  + geom_text(aes(label=percent(aiccw), vjust=-4)))
fossil.bar
```

We can also visualize the estimated evolutionary rates of the trait.  
First create an adjusted version of the plotting function from RPANDA:
```{r}
# Adjust the RPANDA plotting function so we can fix the axes, and do a bunch of plots
plot.fixed_t_env <- function (x, steps = 100, 
                              xlim=c(-10,0), ylim=c(0,1), linecol="red", ...) 
{
  if (is.function(x$model)) {
    fun_temp <- function(x, temp, model, param) {
      rate_fun <- function(x) {
        model(x, temp, param)
      }
      rate <- rate_fun(x)
      return(rate)
    }
  }
  else if (x$model == "EnvExp") {
    fun_temp <- function(x, temp, model, param) {
      sig <- param[1]
      beta <- param[2]
      rate <- (sig * exp(beta * temp(x)))
      return(rate)
    }
  }
  else if (x$model == "EnvLin") {
    fun_temp <- function(x, temp, model, param) {
      sig <- param[1]
      beta <- param[2]
      rate <- sig + (beta - sig) * temp(x)
      return(rate)
    }
  }
  t <- seq(0, x$tot_time, length.out = steps)
  rates <- fun_temp(x = t, temp = x$env_func, model = x$model, 
                    param = x$param)
  plot(-t, rates, type = "l", xlab = "Times", 
       ylab = bquote(paste("Evolutionary rates ",sigma)), 
       xlim=xlim, ylim=ylim, col=linecol, ...)
  results <- list(time_steps = t, rates = rates)
  invisible(results)
}
```

I'll plot just the results from the GRASS-linear model, but you could do this for all.
```{r}
all.aics <- readRDS("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/FossilUncertainty/Data/FINAL_SAMPLED_Model_Fitting_AICCs.RDS")
all.grass.results <- readRDS("/Users/Ian/Desktop/REAL_SAMPLED_GRASS_Model_Fitting_AllResults.RDS")
sub.aic <- filter(all.aics, model=="GRASSlin" & aiccw > 0.5); sub.trees <- sub.aic$tree; print(paste(length(sub.trees), "estimates"))
for(i in sub.trees){plot.fixed_t_env(all.grass.results[[i]]$GRASSlin, 
                                     xlim=c(-12,0), ylim=c(0,0.2), 
                                     linecol=randomcoloR::randomColor(1)); 
  par(new=T)}
```

Also plot the beta value of the correlation between rate and the variable
```{r}
# And plot the beta values for preferred models
col.pal <- brewer.pal(8, "Paired")
beta.values <- filter(all.aics, model %in% c("GRASSexp", "FLUXexp", "GRASSlin", "ENVexp") & aiccw >= 0.3); #beta.values <- filter(beta.values, par<70 & par>-25)
ge_beta <- filter(beta.values, model=="GRASSexp"); ge <- ggplot(ge_beta, aes(x=model, y=par)) + geom_violin(fill=col.pal[[4]]) + theme_classic() + geom_boxplot(fill="white", width=0.075) #+ geom_dotplot(binaxis="y", dotsize=0.1, binwidth=5)
gl_beta <- filter(beta.values, model=="GRASSlin"); gl <- ggplot(gl_beta, aes(x=model, y=par)) + geom_violin(fill=col.pal[[3]]) + theme_classic() + geom_boxplot(fill="white", width=0.075) #+ geom_dotplot(binaxis="y", dotsize=0.1, binwidth=5)
fe_beta <- filter(beta.values, model=="FLUXexp" ); fe <- ggplot(fe_beta, aes(x=model, y=par)) + geom_violin(fill=col.pal[[1]]) + theme_classic() + geom_boxplot(fill="white", width=0.075) #+ geom_dotplot(binaxis="y", dotsize=0.1, binwidth=5)
ee_beta <- filter(beta.values, model=="ENVexp"  ); ee <- ggplot(ee_beta, aes(x=model, y=par)) + geom_violin(fill=col.pal[[2]]) + theme_classic() + geom_boxplot(fill="white", width=0.075) #+ geom_dotplot(binaxis="y", dotsize=0.1, binwidth=5)
gridExtra::grid.arrange(gl, ge, fe, ee, nrow=1)
```

# Comparing Estimated Node Ages Across Dating Schemes

Make a quick function to get the DEPTH of a node (from present), instead of the HEIGHT (from root)
```{r}
MRCA.depth <- function(phy){max(nodeHeights(phy)) - findMRCA(phy, tips=c("Macropus_irma", "Wallabia_bicolor"), type="height")}
```

Prepare trees for comparison of ages
```{r}
# Prepare trees for comparison of ages
max.age.trees <- read.nexus("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/Operators/REAL_Macropodinae_ALL/Alternative_Dating_Schemes/Max_Run_Combined1234.trees"); 
max.ages <- max.age.trees[sample(1:length(max.age.trees),300)]
#max.ages <- lapply(max.age.trees, drop.tip, tip=setdiff(max.age.trees[[1]]$tip.label, overlaps)); class(max.ages) <- "multiPhylo"; 
age.max <- as.data.frame(unlist(lapply(max.ages, MRCA.depth))); 
age.max$tree <- "fixed.max"; colnames(age.max) <- c("age", "tree")

mean.age.trees <- read.nexus("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/Operators/REAL_Macropodinae_ALL/Alternative_Dating_Schemes/Mean_Run_Combined1234.trees"); 
mean.ages <- mean.age.trees[sample(1:length(mean.age.trees),300)]
#mean.ages <- lapply(mean.age.trees, drop.tip, tip=setdiff(mean.age.trees[[1]]$tip.label, overlaps)); class(mean.ages) <- "multiPhylo"
age.mean <- as.data.frame(unlist(lapply(mean.ages, MRCA.depth))); 
age.mean$tree <- "fixed.mean"; colnames(age.mean) <- c("age", "tree")

min.age.trees <- read.nexus("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/Operators/REAL_Macropodinae_ALL/Alternative_Dating_Schemes/Min_Run_Combined1234.trees"); min.ages <- min.age.trees[sample(1:length(min.age.trees),300)]
#min.ages <- lapply(min.age.trees, drop.tip, tip=setdiff(min.age.trees[[1]]$tip.label, overlaps)); class(min.ages) <- "multiPhylo"
age.min <- as.data.frame(unlist(lapply(min.ages, MRCA.depth))); 
age.min$tree <- "fixed.min"; colnames(age.min) <- c("age", "tree")

sampled500.trees <- read.nexus("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/Operators/REAL_Macropodinae_ALL/Run5_Combined123.trees"); 
sampled.ages <- sampled500.trees[sample(1:length(sampled500.trees),300)]
sampled500.trees <- read.nexus("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/Operators/REAL_Macropodinae_ALL/Run5_Combined123.trees"); 
sampled.ages <- sampled500.trees[(length(sampled500.trees)-200):length(sampled500.trees)]
#min.ages <- lapply(min.age.trees, drop.tip, tip=setdiff(min.age.trees[[1]]$tip.label, overlaps)); class(min.ages) <- "multiPhylo"
age.sampled <- as.data.frame(unlist(lapply(sampled.ages, MRCA.depth))); age.sampled$tree <- "est.prior"; colnames(age.sampled) <- c("age", "tree")
```

Combine the ages from different dating schemes, and plot them
```{r}
age.all <- rbind(age.min, age.sampled, age.mean, age.max)
age.all$tree <- factor(age.all$tree, levels=c("fixed.max","est.prior", "fixed.mean", "fixed.min"))
(ggplot(age.all, aes(x=age, fill=tree))
  + geom_density(alpha=0.75, adjust=1.5)
  + theme(axis.text.x=element_text(angle=0, hjust=1), panel.background=element_blank(), legend.position="bottom")
  + scale_fill_manual(values=wes_palette("Zissou1", type="continuous", 4))
  + scale_x_reverse(lim=c(8,0)))
```


# Getting Bayes Factors for Fossil Taxa
### Investigating Fossil Taxa as putative Sampled Ancestors or Terminals
```{r}
prior.trees <- read.nexus("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/Operators/REAL_Macropodinae_ALL/Alternative_Dating_Schemes_PriorOnly/Mean_1/All_Macropodinae_FBD_Relax_PRIOR.trees"); 
prior.trees <- prior.trees[(length(prior.trees)-1000):length(prior.trees)]
post.trees  <- read.nexus("/Users/Ian/Google.Drive/ANU Herp Work/Macropod_Dating/Operators/REAL_Macropodinae_ALL/Alternative_Dating_Schemes/Mean_Run_Combined1234.trees"); 
post.trees <- post.trees[(length(post.trees)-1000):length(post.trees)]


```

Choose which tips you want information for:
```{r}
fossil_taxa <- c("Baringa_nelsonensis","Congruus_congruus",
                 "Kurrabi_mahoneyi","Macropus_pavana")
```
Get the node numbers of the tips
```{r}
nodes <- sapply(fossil_taxa,function(x,y) which(y==x),y=tree$tip.label)
```
Then get the edge lengths for those nodes
```{r}
edge.lengths <- setNames(tree$edge.length[sapply(nodes,
                                               function(x,y) which(y==x),y=tree$edge[,2])],names(nodes))
```
The faster way is to make a function to do this:
```{r}
get_terminal_branchlengths <- function(phy, tipnames){
  ## Get the node numbers of the tips
  nodes <- sapply(tipnames,function(x,y) which(y==x),y=phy$tip.label)
  ## Then get the edge lengths for those nodes
  edge.lengths <- setNames(phy$edge.length[sapply(nodes,
                                                   function(x,y) which(y==x),y=phy$edge[,2])],names(nodes))
  return(edge.lengths)
}
```
Now that we've got the tips and branch lengths, we can compare the posterior to the prior
```{r}
BFSA <- function(prior.phy, posterior.phy, tips){
  post <- lapply(posterior.phy, get_terminal_branchlengths, tipnames=tips); names(post) <- NULL; post <- unlist(post)
  prior <- lapply(prior.phy,    get_terminal_branchlengths, tipnames=tips); names(prior)<- NULL; prior <- unlist(prior)
  
  BFs <- NULL
  for (j in 1:length(tips)){
    curr.tip <- subset(post, names(post)==tips[j]); 
    probSA <- sum(curr.tip<=0); probTIP <- length(curr.tip)-probSA;
    
    curr.tip <- subset(prior, names(prior)==tips[j]);
    priorSA <- sum(curr.tip<=0); priorTIP <- length(curr.tip)-priorSA;
    
    curr.BF <- log((probSA * priorTIP) / (probTIP * priorSA))
    if(is.na(curr.BF)){curr.BF <- 0}
    
    #curr.BF <- log(probSA/(length(curr.tip)-probSA))
    names(curr.BF) <- tips[j]; curr.BF <- round(curr.BF, 2)
    BFs <- append(BFs, curr.BF)
  }
  return(BFs)
}
```
Orient the data appropriately
```{r}
macro_BFs <- BFSA(prior.trees, post.trees, tips=fossil_taxa)

macro_BFs <- as.data.frame(macro_BFs) # make the vector a data frame
macro_BFs[which(macro_BFs$macro_BFs > 5),] <- 5 # change any really big (INF) numbers to 5
macro_BFs[which(macro_BFs$macro_BFs < -5),] <- -5 # change any really small (-INF) numbers to -5

macro_BFs$taxa <- rownames(macro_BFs); # create column with the taxon names
macro_BFs <- macro_BFs[order(macro_BFs$macro_BFs),] # reorder by BF values
# set colors for plotting
macro_BFs$color <- "black";
#macro_BFs[which(macro_BFs$macro_BFs > 1),]$color <- "#3d98d3"
macro_BFs[which(macro_BFs$macro_BFs < -1),]$color <- "#FF7175"

macro_BFs$taxa <- factor(macro_BFs$taxa, levels=c(macro_BFs$taxa)) # set factors for plotting (NOT NECESSARY)
```
Then plot the data
```{r}
ggplot(macro_BFs, aes(x=taxa, y=macro_BFs, label=macro_BFs)) + 
  geom_ribbon(aes(ymin=-1, ymax=+1)) +
  geom_point(stat='identity', size=8, color=macro_BFs$color)  +
  # geom_segment(aes(y = 0, 
  #                  x = taxa, 
  #                  yend = macro_BFs, 
  #                  xend = taxa), 
  #              color = macro_BFs$color) +
  geom_text(color="white", size=2) +
  #labs(title="Bayes Factor Support", subtitle="for Fossil Taxa as Sampled Ancestors") + 
  #ylim(-5, 5) +
  scale_y_continuous(name="log Bayes Factors", limits=c(-5,5), breaks=c(-5:5)) +
  scale_x_discrete(limits = rev(unique(sort(macro_BFs$taxa)))) + # drop this if you want to order it differently
  #theme(panel.background=element_blank()) +
  geom_hline(yintercept=-1) +
  geom_hline(yintercept=1) +
  xlab("Fossil Taxa") +
  #ylab("log Bayes Factors") +
  theme_classic() +
  #theme(axis.text.y=element_blank(), axis.title.y=element_blank()) +
  coord_flip()
```

# Investigating estimated ages of fossils

Create a function to pull the ages of each fossil taxon estimated
```{r}
get.fossil.ages <- function(fossil.tips, trees){
  tree.tables <- lapply(trees, print.tree)
  fossil.tables <- lapply(1:length(tree.tables), function(x) {
    subset(tree.tables[[x]], tree.tables[[x]]$label %in% fossil.tips)
  })
  fossil.ages <- lapply(1:length(fossil.tables), function(x) {
    select(fossil.tables[[x]], label, time_bp)
  })
  final <- bind_rows(fossil.ages)
}
```

Need to create a function called "print.tree" that I borrowed some code from biogeoBEARS
```{r echo=FALSE, eval}
library(BioGeoBEARS)
print.tree <- function (t, printflag = TRUE, relabel_nodes = FALSE, time_bp_digits = 7, 
          add_root_edge = TRUE, get_tipnames = FALSE, fossils_older_than = 0.6) 
{
  if ("node.label" %in% attributes(t)$names == FALSE) {
    rootnum = get_nodenum_structural_root(t)
    new_node_labels = paste("inNode", rootnum:(rootnum + 
                                                 t$Nnode - 1), sep = "")
    t$node.label = new_node_labels
  }
  if (relabel_nodes == TRUE) {
    rootnum = get_nodenum_structural_root(t)
    new_node_labels = paste("inNode", rootnum:(rootnum + 
                                                 t$Nnode - 1), sep = "")
    t$node.label = new_node_labels
  }
  labels = c(t$tip.label, t$node.label)
  ordered_nodenames = get_nodenums(t)
  node.types1 = rep("tip", length(t$tip.label))
  node.types2 = rep("internal", length(t$node.label))
  node.types2[1] = "root"
  node.types = c(node.types1, node.types2)
  parent_branches = get_indices_where_list1_occurs_in_list2(ordered_nodenames, 
                                                            t$edge[, 2])
  brlen_to_parent = t$edge.length[parent_branches]
  parent_nodes = t$edge[, 1][parent_branches]
  daughter_nodes = lapply(ordered_nodenames, get_daughters, 
                          t)
  root_nodenum = get_nodenum_structural_root(t)
  tmpstr = paste("prt(t): root=", root_nodenum, "\n", sep = "")
  #prflag(tmpstr, printflag = printflag)
  levels_for_nodes = unlist(lapply(ordered_nodenames, get_level, 
                                   t))
  hts_at_end_of_branches_aka_at_nodes = t$edge.length
  hts_at_end_of_branches_aka_at_nodes = get_all_node_ages(t)
  h = hts_at_end_of_branches_aka_at_nodes
  times_before_present = get_max_height_tree(t) - h
  edge_ages = t$edge
  edge_ages[, 1] = h[t$edge[, 1]]
  edge_ages[, 2] = h[t$edge[, 2]]
  edge_times_bp = t$edge
  edge_times_bp[, 1] = times_before_present[t$edge[, 1]]
  edge_times_bp[, 2] = times_before_present[t$edge[, 2]]
  if (get_tipnames == TRUE) {
    list_of_clade_members_lists = rep(list(NA), length(ordered_nodenames))
    list_of_clade_members_lists[1:length(t$tip.label)] = t$tip.label
    list_of_clade_members_lists
    nontip_nodenums = (length(t$tip.label) + 1):length(ordered_nodenames)
    if (length(nontip_nodenums) > 1) {
      nontip_nodenames = ordered_nodenames[nontip_nodenums]
      nontip_cladelists = sapply(X = nontip_nodenames, 
                                 FUN = get_all_daughter_tips_of_a_node, t = t)
      nontip_cladelists
      nontip_cladelists_alphabetical = sapply(X = nontip_cladelists, 
                                              FUN = sort)
      nontip_cladelists_alphabetical
      nontip_cladelists_alphabetical_str = sapply(X = nontip_cladelists_alphabetical, 
                                                  FUN = paste, collapse = ",")
      nontip_cladelists_alphabetical_str
      list_of_clade_members_lists[nontip_nodenums] = nontip_cladelists_alphabetical_str
      list_of_clade_members_lists
    }
    else {
      nontip_nodenames = ordered_nodenames[nontip_nodenums]
      nontip_cladelists = sapply(X = nontip_nodenames, 
                                 FUN = get_all_daughter_tips_of_a_node, t = t)
      nontip_cladewords = unlist(sapply(X = nontip_cladelists, 
                                        FUN = strsplit, split = ","))
      nontip_cladelists_alphabetical = sort(nontip_cladewords)
      nontip_cladelists_alphabetical
      nontip_cladelists_alphabetical_str = paste(nontip_cladelists_alphabetical, 
                                                 collapse = ",", sep = "")
      nontip_cladelists_alphabetical_str
      list_of_clade_members_lists[nontip_nodenums] = nontip_cladelists_alphabetical_str
      list_of_clade_members_lists
    }
  }
  fossils = times_before_present > fossils_older_than
  tmpnodenums = (length(t$tip.label) + 1):(length(t$tip.label) + 
                                             t$Nnode)
  fossils[tmpnodenums] = NA
  if (get_tipnames == FALSE) {
    tmpdtf = cbind(1:length(ordered_nodenames), ordered_nodenames, 
                   levels_for_nodes, node.types, parent_branches, brlen_to_parent, 
                   parent_nodes, daughter_nodes, h, round(times_before_present, 
                                                          digits = time_bp_digits), fossils, labels)
    dtf = as.data.frame(tmpdtf, row.names = NULL)
    names(dtf) = c("node", "ord_ndname", "node_lvl", "node.type", 
                   "parent_br", "edge.length", "ancestor", "daughter_nds", 
                   "node_ht", "time_bp", "fossils", "label")
    dtf = unlist_dtf_cols(dtf, printflag = FALSE)
  }
  else {
    tmpdtf = cbind(1:length(ordered_nodenames), ordered_nodenames, 
                   levels_for_nodes, node.types, parent_branches, brlen_to_parent, 
                   parent_nodes, daughter_nodes, h, round(times_before_present, 
                                                          digits = time_bp_digits), fossils, labels, list_of_clade_members_lists)
    dtf = as.data.frame(tmpdtf, row.names = NULL)
    names(dtf) = c("node", "ord_ndname", "node_lvl", "node.type", 
                   "parent_br", "edge.length", "ancestor", "daughter_nds", 
                   "node_ht", "time_bp", "fossils", "label", "tipnames")
    dtf = unlist_dtf_cols(dtf, printflag = FALSE)
  }
  if ((add_root_edge == TRUE) && (!is.null(t$root.edge))) {
    root_row_TF = dtf$node.type == "root"
    root_edge_length = t$root.edge
    dtf$edge.length[root_row_TF] = root_edge_length
    dtf$node_ht = dtf$node_ht + root_edge_length
  }
  #prflag(dtf, printflag = printflag)
  return(dtf)
}

test <- print.tree(post.trees[[1]])
```
Now pull out the info on those fossils
```{r}
my.test <- get.fossil.ages(fossil.tips = fossil_taxa, trees = post.trees)

(ggplot(my.test, aes(x=time_bp, y=label, fill=..x..)) 
  + scale_fill_gradientn(colours=wes_palette("Zissou1"))
  + geom_density_ridges_gradient(scale=1.5)
  + scale_x_reverse()
  + theme_classic())
```

# 
